---
title: "Data Cleaning & Manipulation"
author: "Masha Monakhova"
date: "2023-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r}
library(dplyr)
library(tidyr)
library(readr)
```

## Load our data

```{r}
# load data
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"))

```

The select command from the dplyr package selects **columns*

The filer command filters **rows**

```{r}
## Remove the All and notesRegCode columns from the catch_original data frame

catch_data <- catch_original %>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

# catch_data <- catch_original %>%
#  select(-All, -notesRegCode)

head(catch_data)

glimpse(catch_data)
```

```{r}
# Mutate chinook class from character to numeric
catch_clean <- catch_data %>%
  mutate(Chinook = as.numeric(Chinook))

# search for NA value
i <- which(is.na(catch_clean$Chinook))
i

# located our NA value! It's an uppercase I
catch_data[i, ]

catch_clean <- catch_data %>%
  mutate(Chinook = if.else(Chinook == "I", "1", Chinook)) %>%
  mutate(Chinook = as.integer(Chinook))

```

## Change shape of data using pivot longer and pivot wider

```{r}
# pivot longer practice
catch_long <- catch_clean %>%
  pivot_longer(cols = -c(Region, Year), names_to = "Species", values_to = "catch")

View(catch_long)

# pivot wide practice

catch_wide <- catch_long %>%
  pivot_wider(names_from = Species, values_from = catch)

head(catch_wide)
```

## Practice renaming columns

```{r}
catch_long <- catch_long %>%
  rename(catch_thousands = catch)
```

## Add columns with mutate function

```{r}
catch_long <- catch_long %>%
  mutate(catch = catch_thousands * 1000)

head(catch_long)
```

## Practice group_by and summarize

```{r}
mean_region <- catch_long %>%
  group_by(Region) %>%
  summarize(catch_mean = mean(catch))

head(mean_region)

# summarize count by region

n_region <- catch_long %>%
  group_by(Region) %>%
  summarize(count = n())

head(n_region)
```

## Lesson on filtering

```{r}
ALU_catch <- catch_long %>%
  filter(Region == "ALU")

head(ALU_catch)
```

## Arrange some rows! sad-woo

```{r}
# arrange rows (by default from smallest to largest)
mean_region_arrange <- catch_long %>%
  group_by(Region) %>%
  summarize(mean_catch = mean(catch)) %>%
  arrange(mean_catch)

head(mean_region_arrange)
```
## We're practicing our joins

```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>%
  select(code, mgmtArea)
head(region_defs)
```

## Join data
```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```

## Get rid of "catch_thousands in "catch_long"

```{r}
# catch_long <- catch_long %>%
# select(-catch_thousands)
```

```{r}
region_defs <- region_defs %>%
  rename(Region = code, Region_name = mgmtArea)
catch_joined <- left_join(catch_long, region_defs, by = "Region")
head(catch_joined)
```

```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-219",
                                "MAI-039"))
sites_df %>%
  separate(site, c("island", "site_number"), "_")
```
 ## Unite
```{r}
dates_df <- data.frame(year = c("1930",
                                "1930",
                                "1930"),
                       month = c("12",
                                "12",
                                "12"),
                       day = c("14",
                               "15",
                               "16"))

dates_df %>% 
  unite(date, year, month, day, sep = "-")
```
 
