---
title: "Practice RMarkdown"
output: html_document
date: "2023-02-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
```

## Introduction

This data set contains diverse measurements of seawater chemistry along with supporting conductivity, temperature, and depth (CTD) data from the Arctic Ocean near the North Pole.

The data was downloaded from [here] (https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)


## Read Data

```{r }
bg_chem <- read_csv("data/BGchem2008data.csv")
```
```{r exploration, echo=FALSE, eval=FALSE}
colnames(bg_chem)

str(bg_chem)

head(bg_chem)
```


## Analysis



## Calculate summary statistics 


```{r summary}
nitrate <- mean(bg_chem$NO3)

nitrite <- mean(bg_chem$NO2)

amm <- mean(bg_chem$NH4)

phos <- mean(bg_chem$P)
```

## Calculate mean Redfield ratio 

```{r mean}
ratio <- (nitrate +nitrite + amm)/phos
```

The Redfield ratio in this dataset is approximately `r round(ratio)`

## Plot Redfield ratio

```{r plot}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

